Assignment 1: Data and Descriptive

### 2023-11-22 MSB104_GR_1_Final_markdown.qmd
Assignment 3 :)

---
title: "Regional Inequalities and economic growth; cross section and times series analysis"
author: "Anine Therese Karlsen & Mona Lisa Jones"
format: pdf
abstract: "This assignment aims to acquire, process, and analyze sub-national GDP and population data for a subset of European countries. Firstly (assignment 1), Calculateing GDP per capita and exploreing regional inequity using various descriptive statistics and visualizations.Secondly (assignment 2), examining the relationship between regional development and inequality, employing cross sectional estimation techniques for the year 2010.Lastly exploring alternative functional forms and implementing panel estimation to analzise the relationship between regional development inequality"
editor: visual
bibliography: [Regional_inequality_23.bib, econometrics.bib]
page-layout: full
number-sections: true
papersize: a4
csl: apa.csl
---

@@

Furthermore, exploring alternative functional forms and implementing panel estimation to analzise the relationship between regional development inequality (***Assignment_3***).

Lastly, it is worth noting potential limitations may arise due to the use of ChatGPT and ChatGPT PDF, which have been utilized as tools to gain understanding and extract knowledge from academic papers, as well as to explain econometric terms and definitions. The text generated by these tools has been cross-checked to the best of our knowledge with our textbook and Andre Seidel, lecturer and author of 'Regional Inequality, Convergence, and Its Determinants -- A View from Outer Space' (2017). Despite these precautions, there may have been instances where incorrect information was provided, and some errors could have been overlooked. The text has been meticulously rewritten to fit our style of writing, ensuring consistency and accuracy throughout our study.

@@

Why some places grow and prosper is a fundamental question in social science research. This question motivated Adam Smith's work, *An Inquiry into the Nature and Causes of the Wealth of Nations,* (Smith [1863](https://link-springer-com.galanga.hvl.no/article/10.1007/s11187-014-9574-4#ref-CR32 "Smith, A. (1863). An inquiry into the nature and causes of the wealth of nations. Edinburgh: A. & C Black.")) and has been a major influence of the *new economic geography* [@feldman2014]*. Furthermore,* regional growth and inequality have been important topics of research in Europe for many years. In recent decades, the European Union implemented policies aimed at reducing regional disparities and promoting economic growth across its member states. Furthermore, despite these efforts, significant disparities in income and economic development persist across regions in Europe [@lessmann2017].

@@

Regional growth in Europe exhibits a complex pattern, influenced significantly by factors such as geography, human capital, and economic policies. Lessmann and Seidel's research offers insights into these dynamics, highlighting the persistence of income disparities, across regions despite overarching economic advancement [@lessmann2017]. Wealthier nations like Denmark tend to experience quicker regional convergence, a phenomenon supported by Gennaioli et al.'s findings [@gennaioli2014].

@@

EU membership provided Slovakia and Hungary with resources for economic recovery, while Western European countries, including Denmark, Portugal, and France, likely experienced significant economic declines due to the financial crisis, a pattern possibly mirrored in the recent COVID-19 pandemic. In particular Portugal as the crisis hit them harder. These complexities highlight the importance of a detailed analysis to understand regional economic growth and inequality in challenging times.

@@

@lessmann2017 regional inequality study address the importance of studying these aspects. Its potential consequences, such as political tensions that can undermine social and political stability [@lessmann2017]. The paper also discusses the relationship between regional inequality and personal income inequality and conflict risk.

@@

Urbanization emerges as a crucial determinant, where bustling, populous around urban centers usually enjoy elevated income levels, contrasting starkly with the often-stark disparities witnessed in more rural and secluded areas [@lessmann2017].

@@

[@lessmann2017] also discusses the determinant of Education in the context of human capital. The study mentions that human capital is the most important determinant of differences in regional development within countries. The study also notes that human capital creates positive externalizes, which can spill over regional boundaries. Moreover, the quality of human capital within countries, as measured by the secondary-school enrollment rate, can decrease regional inequality by facilitating regional spillovers and convergence. Lastly, promoting internal migration.

@@

Furthermore, @lessmann2017 provides us with insights into the inverted U-shaped relationship between regional inequality and the level of economic development in different country groups [@lessmann2017]. The inverted U-shape also referred to as Kuznets Curve suggests that as an economy develops, regional inequality initially rises, reaches a peak, and then starts to decline.

@@

@iammarino2019 discusses regional inequalities in European countries. It presents evidence that inter-regional inequality has been increasing in the European Union since the turn of the millennium. The authors argue that this is due to the existence of several groups of regional economies in Europe that are structurally very different from one another [@iammarino2019]. Additionally, the authors argue that there are some countries in the EU that are more evenly developed than others, and that a map of under performance or over-performance means less in a high-income but evenly developed country, such as in Denmark in our case [@iammarino2019]. On the other hand, France has been a subject of discussion and policy making for many years, regarding its well-known pockets of extreme wealth in regions such as Cap d'Antibes.

@@

# Data and Descriptive

Trough Eurostat, we download the datasets nama_10r_3gdp , demo_r_pjanggr3, demo_r_mlifexp, tran_r_net and edat_lfse_04 as csv files. Selecting our countries of study and timeperiod. Selecting the required NUTS level and specified the data to be in million Euro.

@@

This data set (edat_lfse_04) contains population by educational attainment level, sex and NUTS 2 regions (%). Including data on the highest level of education successfully completed by the individuals of a given population. Furthermore, data on young people neither in employment nor in education and training -- NEET, early leavers from education and training and the labor status of young people by years since completion of highest level of education. Calculated as annual averages of quarterly EU Labor Force Survey data (EU-LFS).

@@

@@

# Cross sectional estimates

@@

The next step is making a data frame of NUTS2 2010 and grouping by NUTS0 country level variables, as well as allocating NUTS2 to selected countries. Lastly presenting data in **table XXX** showing a count of NUTS2 regions within each top-level region or country in the 2010 data.

@@

**Model assumptions**

@@

**Model specification**

@@

**Model Diagnostics**

@@

**Ordinary Least Squares (OLS) Estimation**

@@

### **Assumptions**

@@

**Visualization**

@@

### I. Data Acquisition

@@

### II. Multiple Linear Regression Model

@@

**Understanding the Coefficients**

@@

### III. Model Interpretation

@@

# Alternative functional forms and panel estimation

By exploring our pervious estimates in the relationship between regional development on regional inequality we will estimate the log of GDP per capita in a liner regression model, defining the variables as new determinants on the effects in regional economic development using logarithim of GDP per capita. Using dummy variables to determine if the effect is significantly different in different subset of our countries.

## Part 3A Testing Development effects across subsets

## Part 3B Exploring Alternative Functional Forms

### **The cuznets cuve**

**Logarithmic transformation**

The logarithmic transformation for all the variables allows the coefficients to be interpreted as elasticities, which measure the percentage change in the dependent variable associated with a one percent change in the independent variable [@lessmann2017]. The use of logarithmic transformations is a common practice in econometrics to address issues such as heteroscedasticity and nonlinearity in the data. By using logarithmic transformations, we are able to estimate the relationship between regional inequality and development in a more robust and accurate way [@lessmann2017].

**The quadratic term**

Furthermore the quadratic term in the regression model is used to investigate the relationship between regional inequality and development [@lessmann2017]. The results suggest an inverted U-shaped relationship between income and inequality, which implies that the relationship between inequality and development looks N-shaped instead of U-shaped [@lessmann2017].

**The cubic function**

The cubic function in the regression model is used to investigate the relationship between regional inequality and development. @lessmann2017 suggest that regional inequalities increase again at very high levels of development, which implies that the relationship between inequality and development looks N-shaped instead of U-shaped [@lessmann2017]. The inclusion of a cubic term in a regression model allows the analysis of more complex dynamics, such as increasing or decreasing marginal effects [@wooldridge2020]. However, interpreting these models can be more challenging, as the effect of a one-unit change in the predictor variable on the dependent variable is no longer constant and depends on the level of the predictor variable.

## Part 3C Hetroskedasticity Testing and causality Discussion

**Causality vs causation**

While correlation indicates a relationship between two variables, it doesn't imply causation. In econometrics, it's a common pitfall to interpret correlation as causality. Panel data helps address this by allowing researchers to control for time-invariant factors and observe changes over time, which can provide stronger evidence for causal relationships.

Panel data analysis helps address endogeneity issues, such as omitted variable bias, which can lead to spurious correlation. By tracking the same units over time, it's possible to differentiate whether a relationship is genuinely causal or merely a correlation coinciding in time.

Overall, panel data analysis is a vital tool in econometrics for distinguishing between causality and correlation. By allowing for control over unobserved heterogeneity and providing a dynamic view of data over time, it offers a more nuanced and accurate understanding of the relationships between variables. This is crucial in econometric research, where the distinction between causality and mere correlation may have significant implications.

**Fixed-Effects and Random-Effects Models** in panel data analysis help in controlling for unobserved heterogeneity, thereby providing a clearer picture of whether there's a causal relationship.

## Part 3D Panel Estimates

**Fixed effect models**

Panel data estimation is a method used in econometrics to analyse data that involves observations over multiple time periods. In the context of panel data estimation specifics of fixed-effects and random effect are used. [@lessmann2017] agrues that the fixed-effects model is a reasonable approach when the differences between countries (or regions) can be viewed as parametric shifts of the regression function. However, the random-effects model allows for time-invariant unobserved heterogeneity across regions and is more appropriate when the fixed-effects model is too restrictive [@lessmann2017].

**Random-Effects Models**

@lessmann2017 mention the use of a random-effects model to investigate the determinants of within-country changes in inequality. The random-effects model controls for several country-level fixed factors (national income, number of regions, and area) and fixed effects for various country groups [@lessmann2017]. The advantage of the random-effects model is that the expected value of the country-specific effect is zero, which means that there is no need to apply any arbitrary data imputation procedure for the missing intercepts [@lessmann2017]. However, this approach may come at the cost of founding the predictions on a slightly biased coefficient [@lessmann2017]. The text also notes that the major coefficient of interest is not sensitive to applying either a fixed-effects model or a random-effects model with additional country and region information [@lessmann2017].

>>>>

---
title: "Regional Inequalities and economic growth; cross section and times series analysis"
author: "Anine Therese Karlsen & Mona Lisa Jones"
format: pdf
abstract: "This assignment aims to acquire, process, and analyze sub-national GDP and population data for a subset of European countries. Firstly (assignment 1), Calculateing GDP per capita and exploreing regional inequity using various descriptive statistics and visualizations.Secondly (assignment 2), examining the relationship between regional development and inequality, employing cross sectional estimation techniques for the year 2010."
editor: visual
bibliography: [Regional_inequality_23.bib, econometrics.bib]
page-layout: full
number-sections: true
papersize: a4
csl: apa.csl
---

@@

Lastly, potential limitations may arise due to the use of ChatGPT and ChatGPT PDF, which have been utilized as tools to gain understanding and extract knowledge from academic papers, as well as to explain econometric terms and definitions. The text generated by these tools has been cross-checked to the best of our knowledge with our textbook and Andre Seidel, lecturer and author of 'Regional Inequality, Convergence, and Its Determinants -- A View from Outer Space' (2017). Despite these precautions, there may have been instances where incorrect information was provided, and some errors could have been overlooked. The text has been meticulously rewritten to fit our style of writing, ensuring consistency and accuracy throughout our study.

@@

Why some places grow and prosper is a fundamental question in social science research. This question motivated Adam Smith's work, *An Inquiry into the Nature and Causes of the Wealth of Nations,* (Smith [1863](https://link-springer-com.galanga.hvl.no/article/10.1007/s11187-014-9574-4#ref-CR32 "Smith, A. (1863). An inquiry into the nature and causes of the wealth of nations. Edinburgh: A. & C Black.")) and has been a major influence of the *new economic geography* [@feldman2014]*. Furthermore,* regional growth and inequality have been important topics of research in Europe for many years. In recent decades, the European Union has implemented policies aimed at reducing regional disparities and promoting economic growth across its member states. Furthermore, despite these efforts, significant disparities in income and economic development persist across regions in Europe [@lessmann2017].

@@

Regional growth in Europe exhibits a complex pattern, influenced significantly by factors such as geography, human capital, and economic policies. Lessmann and Seidel's research offers insights into these dynamics, highlighting the persistence of income disparities, across regions despite overarching economic advancement [@lessmann2017]. Wealthier nations like Denmark tend to experience quicker regional convergence, a phenomenon supported by Gennaioli et al.'s findings [@gennaioli2014]. However, this is not uniformly experienced across all European countries, with Eastern European nations such as Hungary and Slovakia we may expect different growth trajectories, particularly after their EU accession.

@@

EU membership provided Slovakia and Hungary with resources for economic recovery, while Western European countries, including Denmark, Portugal, and France, likely experienced significant economic declines due to the financial crisis, a pattern possibly mirrored in the recent COVID-19 pandemic. These complexities highlight the importance of a detailed analysis to understand regional economic growth and inequality in challenging times.

@@

@lessmann2017 regional inequality study address the importance of studying regional inequality. Its potential consequences, such as political tensions that can undermine social and political stability [@lessmann2017]. The paper also discusses the relationship between regional inequality and personal income inequality and conflict risk.

@@

Urbanization emerges as a crucial determinant, where bustling, populous around urban centres usually enjoy elevated income levels, contrasting starkly with the often-stark disparities witnessed in more rural and secluded areas [@lessmann2017].

@@

[@lessmann2017] discusses the determinant of Education in the context of human capital. The study mentions that human capital is the most important determinant of differences in regional development within countries. The study also notes that human capital creates positive externalities, which can spill over regional boundaries. Moreover, the quality of human capital within countries, as measured by the secondary-school enrolment rate, can decrease regional inequality by facilitating regional spillovers and convergence. Lastly, promoting internal migration.

@@

Furthermore, [@lessmann2017] provides us with insights into the inverted U-shaped relationship between regional inequality and the level of economic development in different country groups [@lessmann2017]. The inverted U-shape also referred to as Kuznets Curve suggests that as an economy develops, regional inequality initially rises, reaches a peak, and then starts to decline.

@@

@iammarino2019 discusses regional inequalities in European countries. It presents evidence that inter-regional inequality has been increasing in the European Union since the turn of the millennium. The authors argue that this is due to the existence of several groups of regional economies in Europe that are structurally very different from one another [@iammarino2019]. Additionally, the authors argue that there are some countries in the EU that are more evenly developed than others, and that a map of underperformance or over-performance means less in a high-income but evenly developed country, such as in Denmark in our case [@iammarino2019]. On the other hand, France has been a subject of discussion and policy making for many years, regarding its well-known pockets of extreme wealth in regions such as Cap d'Antibes.

@@

# Data

Trough Eurostat, we download the datasets nama_10r_3gdp , demo_r_pjanggr3, demo_r_mlifexp, tran_r_net and edat_lfse_04 as csv files and filted the data by our preferences before downloading it. Furthermore, we filtered the dataset by choosing the years 2000 to 2020. Then, we selected the NUTS requiered levels region for Portugal, France, Hungary, Slovakia and Denmark. Finally, we specified the data to be in million Euro.

@@

This data set (edat_lfse_04) contains population by educational attainment level, sex and NUTS 2 regions (%). Including data on the highest level of education successfully completed by the individuals of a given population. Furthermore, data on young people neither in employment nor in education and training -- NEET, early leavers from education and training and the labour status of young people by years since completion of highest level of education. Calculated as annual averages of quarterly EU Labour Force Survey data (EU-LFS).

@@

# Econometric Approach

@@

@@

The next step is making a data frame of NUTS2 2010 and grouping by NUTS0 country level variables, as well as allocating NUTS2 to selected countries. Lastly presenting data in tables showing a count of NUTS2 regions within each top-level region or country in the 2010 data.

@@

#### **Model assumptions**

@@

#### **Model specification**

@@

#### **Model Diagnostics**

@@

### **Ordinary Least Squares (OLS) Estimation**

@@

#### **Assumptions**

@@

### **Visualization**

@@

### Data Acquisition

@@

### Multiple Linear Regression Model

@@

#### **Understanding the Coefficients**

@@

### Model Interpretation

@@

### 2023-11-22 assignment_1.qmd
Assignment 3 :)

library(tidyverse)
library(tinytex)
library(ggplot2)
library(readxl)
library(magrittr)
library(vtable)
library(lubridate)
library(dineq)
library(huxtable)
```

@@

@@

We've made two plots that can help us understand the relationship between GDP per capita and the Gini coefficient. The first one is a plot that shows us residuals vs fitted values, which can help us check the homoscedasticity assumption of a linear regression model. The residuals should be randomly scattered around the horisontal 0 line, something that can inducate that the variances of the error terms are constant. In our plot, the residuals are in some extent randomly distributed, and there is also no clear pattern; this suggest that there is likely no significant issues with heteroscedasticity or non-linearity.

The other plot - normal Q-Q plot - is used to assess if the residuals of the linear model are normally distributed. In our plot, most of the points follow the line closely, suggesting that the residuals are normally distributed.

@@

The numbers we gathered earlier, as well as these plots, can help us explain and assert whether the classical OLS assumptions holds for our model:

-   Linearity:

    -

-   Independence:

    -

-   Normality:

    -

-   Homoscedasticity:

    -

-   Multicollinearity:

    -

@@

## Conclusion

>>>>

library(tidyverse)
library(tinytex)
library(ggplot2)
library(readxl)
library(magrittr)
library(vtable)
library(lubridate)
library(dineq)
library(huxtable)
library(sandwich)
library(lmtest)
```

@@

**Scale-Location Plot** is another way to check for homoscedasticity. The spread of the residuals should be roughly the same across all levels of the fitted values. If the spread increases or decreases with the fitted values, it suggests heteroscedasticity.

**Residuals vs. Leverage Plot** to identify influential observations that might unduly influence the regression. Points with high leverage or large residuals (outliers) can be identified. Cook's distance lines may be added to help identify points that have a large influence on the model.

@@

We've made two plots that can help us understand the relationship between GDP per capita and the Gini coefficient, and that together with the regression statistics can help with discussing if the classical OLS assumptions hold for the model. The first one is a plot that shows us residuals vs fitted values, which can help us check the homoscedasticity assumption of a linear regression model. The residuals should be randomly scattered around the horisontal 0 line, something that can inducate that the variances of the error terms are constant. In our plot, the residuals are in some extent randomly distributed, and there is also no clear pattern; this suggest that there is likely no significant issues with heteroscedasticity or non-linearity.

The other plot - normal Q-Q plot - is used to assess if the residuals of the linear model are normally distributed. In our plot, most of the points follow the line closely, suggesting that the residuals are normally distributed. There are however some outliers in the tails, that suggest some variation from normality.

@@

@@

For model 1, the adjusted R-squared is 0.033, indicating that the model explains around 3% of the variability in the gini coefficient. Since this is relatively low, it suggests that model 1 is not suitable in explaining the variance in gini.

# Assignment 3: Alternative Functional Forms and Panel Estimates

## Part A: Testing Development Effects Across Subsets

### Subset Analysis:

To begin the third assignment, we will divide the data into different subsets. We have chosen to divide the countries based on size and population; we'll therefore have one subset with France, and one subset with the rest of the countries. We have several reasons for the dividing the countries like this:

1.  France is one of the largest countries in Europe, with a great economic and political influence. The larger size and population of France may lead to different economic dyamics compared to the other smaller countries.

2.  Larger countries might have a greater diversity of economic activities, wich affects the distribution of wealth and income differentely than in smaller countries with less diverse economies (source). We can see this by looking at how some of the NUTS 2 regions in France differ far more than in other countries.

3.  The size and population can influence public policy, like social welfare, which impacts income distribution and inequality (source).

4.  Larger countries often have larger internal markets, which can affect economic development and income distribution differently compared to smaller economies, where ecternal trade might be more crucial.

| Country  | Population (in million) | Size (km\^2) | NUTS 2 observations |
|----------|-------------------------|--------------|---------------------|
| France   | 67,75                   | 551 695      |                     |
| Denmark  | 5,85                    | 42 952       |                     |
| Slovakia | 5,4                     | 49 000       |                     |
| Hungary  | 9,71                    | 93 000       |                     |
| Portugal | 10,33                   | 92 000       |                     |

# Dataframe with just france
gini_nuts_2_FR <- gini_nuts_2 |> 
  filter(id_nuts0 == "FR", gini_n2 > 0)

# Dataframe with the other countries
gini_nuts_2_C <- gini_nuts_2 |> 
  filter(id_nuts0 != "FR", gini_n2 > 0)
```

# Descriptive analysis
summary(gini_nuts_2_FR)
summary(gini_nuts_2_C)
```

# Regression analysis
# France
lmFR <- lm(gini_n2 ~ log(GDP_per_capita), data = gini_nuts_2_FR)
# Others
lmC <- lm(gini_n2 ~ log(GDP_per_capita), data = gini_nuts_2_C)

huxreg(France = lmFR,
       Others = lmC,
       statistics = c("r.squared","adj.r.squared", "statistic", "p.value"))
```

# Comparative analysis
# Welch two sample t-test
t.test <- t.test(gini_nuts_2_FR$gini_n2, gini_nuts_2_C$gini_n2)
print(t.test)
```

### Subset Analysis Discussion:

In order to discuss the differences between the two subsets, we have done different analysis: descriptive, regression and comparative.

France has a mean gini coefficient of 0.06960, while the mean gini coefficient of the other countries is 0.062308. France has a slightly higher gini, meaning that they on average have a higher amount of economic inequality compared to the rest. The mean of GDP per capita is also higher in france (2.750 \> 1.743), suggesting that they also on average have a higher level of economic development (de har både høyere økonomisk ulikhet og økonomisk utvikling, hva er grunnen til dette? Kan se dette på regression analysis!! r quared har høy positiv korrelasjon).

Looking at the regression analysis of France, we can see that the intercept is -0.034 with a p-value that is lower than 0.001, suggesting that the intercept is different from zero. The coefficient for log of GDP per capita is 0.110 with a p-value lower than 0.001; meaning that as the log of GDP per capita increase, the gini coefficient increase significantly, which suggest that a higher GDP per capita also gives us greater income inequality. Laslty, the R-squared is 0.490, meaning that around 49% of the variance in the gini coefficient can be explained by GDP per capita. *This is a relatively high value,* and suggest a strong model fit.

## Part B: Exploring Alternative Functional Forms

### Functional Form Exploration:

```{r}

# Logarithmic model
log_model <- lm(gini_n2 ~ log(GDP_per_capita), data = gini_filtered)

# Quadratic model
quad_model <- lm(gini_n2 ~ GDP_per_capita + I(GDP_per_capita^2), data = gini_filtered)

# Cubic model
cubic_model <- lm(gini_n2 ~ GDP_per_capita + I(GDP_per_capita^2) + I(GDP_per_capita^3), data = gini_filtered)

# Check the summary of each model
summary(log_model)
```

```{r}
summary(quad_model)
```

```{r}
summary(cubic_model)
```

###

### Estimation and Visualization:

```{r}
# Quadratic model

# Create a new data frame for plotting predicted values
plot_data <- data.frame(GDP_per_capita = seq(min(gini_filtered$GDP_per_capita), max(gini_filtered$GDP_per_capita), length.out = 250))
plot_data$gini_pred <- predict(quad_model, newdata = plot_data)

# Plotting
ggplot(gini_filtered, aes(x = GDP_per_capita, y = gini_n2)) +
  geom_point(alpha = 0.5) +  # Plot the actual data points
  geom_line(data = plot_data, aes(x = GDP_per_capita, y = gini_pred), color = 'blue') +  # Add the regression line
  labs(title = "Quadratic Model of GDP per Capita and Gini Coefficient",
       x = "GDP per Capita",
       y = "Gini Coefficient") +
  theme_minimal()
```

```{r}
# Logarithmic model

# Create a new data frame for plotting predicted values
plot_data <- data.frame(GDP_per_capita = seq(min(gini_filtered$GDP_per_capita), max(gini_filtered$GDP_per_capita), length.out = 250))
plot_data$gini_pred <- predict(log_model, newdata = plot_data)

# Plotting
ggplot(gini_filtered, aes(x = GDP_per_capita, y = gini_n2)) +
  geom_point(alpha = 0.5) +  # Plot the actual data points
  geom_line(data = plot_data, aes(x = GDP_per_capita, y = gini_pred), color = 'red') +  # Add the regression line
  scale_x_log10() +  # Logarithmic scale for the x-axis
  labs(title = "Logarithmic Model of GDP per Capita and Gini Coefficient",
       x = "GDP per Capita (log scale)",
       y = "Gini Coefficient") +
  theme_minimal()

```

```{r}

# Create a new data frame for plotting predicted values
plot_data <- data.frame(GDP_per_capita = seq(min(gini_filtered$GDP_per_capita), max(gini_filtered$GDP_per_capita), length.out = 250))
plot_data$gini_pred <- predict(cubic_model, newdata = plot_data)

# Plotting
ggplot(gini_filtered, aes(x = GDP_per_capita, y = gini_n2)) +
  geom_point(alpha = 0.5) +  # Plot the actual data points
  geom_line(data = plot_data, aes(x = GDP_per_capita, y = gini_pred), color = 'blue') +  # Add the regression line
  labs(title = "Cubic Model of GDP per Capita and Gini Coefficient",
       x = "GDP per Capita",
       y = "Gini Coefficient") +
  theme_minimal()

```

```{r}
# Calculate residuals
residuals_quad <- gini_filtered$gini_n2 - predict(quad_model, gini_filtered)
residuals_cubic <- gini_filtered$gini_n2 - predict(cubic_model, gini_filtered)
```

```{r}
# Create a dataframe for plotting residuals
residuals_data <- data.frame(GDP_per_capita = gini_filtered$GDP_per_capita, 
                             Residuals_Quad = residuals_quad, 
                             Residuals_Cubic = residuals_cubic)

# Plotting residuals
ggplot(residuals_data) +
  geom_line(aes(x = GDP_per_capita, y = Residuals_Quad, color = "Quadratic")) +
  geom_line(aes(x = GDP_per_capita, y = Residuals_Cubic, color = "Cubic")) +
  labs(title = "Comparison of Residuals",
       x = "GDP per Capita",
       y = "Residuals") +
  theme_minimal() +
  scale_color_manual(values = c("Quadratic" = "blue", "Cubic" = "red"))

```

```{r}
sd(residuals_quad)
sd(residuals_cubic)
mean(residuals_quad^2)
mean(residuals_cubic^2)
```

### Results Interpretation:

## Part C: Heteroskedasticity Testing and Causality Discussion

### Heteroskedasticity Testing:

```{r}
lm_01 <- lm(gini_n2 ~ GDP_per_capita, data = gini_filtered)
```

```{r}

# Obtain fitted values and residuals from the model
fitted_values <- fitted(lm_01)
residuals <- resid(lm_01)

# Create the plot
plot(fitted_values, residuals, 
     xlab = "Fitted Values", 
     ylab = "Residuals", 
     main = "Residuals vs Fitted Values for Linear Model",
     pch = 20)
abline(h = 0, col = "blue", lwd = 2) # Add a horizontal line at 0

```

```{r}

# Breusch-Pagan test for the original linear model
bptest_linear <- bptest(lm_01)

# Breusch-Pagan test for the logarithmic model
bptest_log <- bptest(log_model)

# Breusch-Pagan test for the quadratic model
bptest_quad <- bptest(quad_model)

# Output the test results
print(bptest_linear)
print(bptest_log)
print(bptest_quad)
```

The Breusch-Pagan test is used to detect heteroskedasticity in a regression model. The test works by checking if the variances of the errors from the regression are dependent on the values of the independent variables, which would violate one of the key assumptions of ordinary least squares (OLS) regression.

```{r}

# Calculate robust standard errors for the original linear model
coeftest_lm_01 <- coeftest(lm_01, vcov = vcovHC(lm_01, type = "HC"))

# Calculate robust standard errors for the logarithmic model
coeftest_log_model <- coeftest(log_model, vcov = vcovHC(log_model, type = "HC"))

# Calculate robust standard errors for the quadratic model
coeftest_quad_model <- coeftest(quad_model, vcov = vcovHC(quad_model, type = "HC"))

# Print out the results
print(coeftest_lm_01)
print(coeftest_log_model)
print(coeftest_quad_model)

```

### Causality Discussion:

## Part D: Panel Estimates

### Panel Estimation Task:

### Panel Estimation Analysis:

### Panel Estimation Discussion:
